---
title: "Quarantine"
output: rmarkdown::html_vignette
bibliography: references.bib
link-citations: true
vignette: >
  %\VignetteIndexEntry{Quarantine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
```

## Understanding quarantine in the `{ringbp}` model

The epidemiological model implemented in the `{ringbp}` R package contains the option of quarantine. This can be _turned on_ or _off_ (by default it is off, see `intervention_opts()`). 

Quarantine in the model has a relatively simple interpretation and implementation. 

_Interpretation_: If quarantine is turned off, then assuming an infectee is not asymptomatic nor missed by contact tracing, they are isolated at their symptom onset time if after their infector's isolation time, otherwise at the earlier of the infector and infectee isolation times. So an infectee needs to be symptomatic to isolate when quarantine is not active. If quarantine is turned on, the infectee is quarantined/isolated at the earlier of the infector's or infectee's isolation time. The difference when quarantine is active is infectees can be isolated at their infector's isolation time independent of whether they are showing symptoms. Therefore, when quarantine is active an infectee is isolated at the same time or earlier than if quarantine was not active, and never later.

_Implementation_: The time at which an infectee is isolated is sampled in `outbreak_step()`. Expand the link below to see the code in `outbreak_step()` that does the sampling depending on if quarantine is turned on or off (i.e. `TRUE` or `FALSE`). 

<details>
<summary>
Code snippet from `outbreak_step()`
</summary>
```r
prob_samples[, isolated_time := {
    ref_time <- onset + delays$onset_to_isolation(.N)
    fcase(
      # asymptomatic: never isolated
      asymptomatic, Inf,
      # symptomatic, false-negative test: never isolated
      !test_positive, Inf,
      # symptomatic, test-positive, not traced: isolated at symptom onset + delay
      !traced, ref_time,
      # symptomatic, test-positive, traced, quarantine:
      # earliest of infector / infectee isolation time
      rep(interventions$quarantine, .N), pmin(ref_time, infector_isolation_time),
      # symptomatic, test-positive, traced, no quarantine:
      # earliest of symptom onset + delay / infector isolation time
      default = pmin(ref_time, pmax(onset, infector_isolation_time))
    )
  }]
```
</details>

This is the only part of the `{ringbp}` model that quarantine affects.

::: {.alert .alert-info}

For a visual schematic of the interventions in the `{ringbp}` model, see the [`{ringbp}` Model Description vignette](ringbp-model.html).

:::

---

In the rest of this vignette we explore the behaviour of quarantine in the `{ringbp}` model, and investigate the effect on outbreak transmission dynamics under different model parameterisations.

```{r setup}
library(ringbp)
library(data.table)
library(tinyplot)
library(knitr)
```

## Demonstrating the interaction of quarantine with asymptomatics and contact tracing

In the _interpretation_ given above, it stated that both the active and inactive quarantine scenarios require assuming an infectee is not asymptomatic nor missed by contact tracing. This is because if an infectee is asymptomatic they cannot be isolated or quarantined, thus, there is no impact of quarantine. Secondly, isolating or quarantining an infectee requires tracing them; meaning if they are missed by contact tracing they are isolated after a delay from symptom onset, irrespective of whether quarantine is active. 

Here we briefly demonstrate how asymptomatic and missed infectees reduce the effect of quarantine to zero. In a scenario where all cases are asymptomatic the outbreak is identical whether quarantine is active or inactive.

```{r compare-asym}
offspring <- offspring_opts(
  community = \(n) rnbinom(n = n, mu = 2.5, size = 1),
  isolated = \(n) rnbinom(n = n, mu = 0, size = 1)
)
delays <- delay_opts(
  incubation_period = \(n) rweibull(n = n, shape = 2.32, scale = 6.49),
  onset_to_isolation = \(n) rweibull(n = n, shape = 2.5, scale = 5)
)
sim <- sim_opts(
  cap_max_days = 365,
  cap_cases = 5000
)

nsim <- 100
initial_cases <- 5

set.seed(1)
outbreak_no_Q <- scenario_sim(
  n = nsim,
  initial_cases = initial_cases,
  offspring = offspring,
  delays = delays,
  event_probs = event_prob_opts(
    asymptomatic = 1,
    presymptomatic_transmission = 0.15,
    symptomatic_traced = 0.5
  ),
  interventions = intervention_opts(quarantine = FALSE),
  sim = sim
)

set.seed(1)
outbreak_Q <- scenario_sim(
  n = nsim,
  initial_cases = initial_cases,
  offspring = offspring,
  delays = delays,
  event_probs = event_prob_opts(
    asymptomatic = 1,
    presymptomatic_transmission = 0.15,
    symptomatic_traced = 0.5
  ),
  interventions = intervention_opts(quarantine = TRUE),
  sim = sim
)

identical(outbreak_no_Q, outbreak_Q)
```

The same is true, that quarantine has no effect on disease transmission, if we assume that there is no contact tracing (contact ascertainment = 0%) and infectees isolate after an onset-to-isolation delay.

```{r compare-ct}
set.seed(1)
outbreak_no_Q <- scenario_sim(
  n = nsim,
  initial_cases = initial_cases,
  offspring = offspring,
  delays = delays,
  event_probs = event_prob_opts(
    asymptomatic = 0.5,
    presymptomatic_transmission = 0.15,
    symptomatic_traced = 0
  ),
  interventions = intervention_opts(quarantine = FALSE),
  sim = sim
)

set.seed(1)
outbreak_Q <- scenario_sim(
  n = nsim,
  initial_cases = initial_cases,
  offspring = offspring,
  delays = delays,
  event_probs = event_prob_opts(
    asymptomatic = 0.5,
    presymptomatic_transmission = 0.15,
    symptomatic_traced = 0
  ),
  interventions = intervention_opts(quarantine = TRUE),
  sim = sim
)

identical(outbreak_no_Q, outbreak_Q)
```

To conclude this section we finish by showing that if some cases are symptomatic, here we assume 50% of cases develop symptoms, and if some contacts are traced, we assume 50% of contacts are traced, then the quarantine influences the outbreak and the resulting outbreak data is different.

```{r compare-ct-asym}
set.seed(1)
outbreak_no_Q <- scenario_sim(
  n = nsim,
  initial_cases = initial_cases,
  offspring = offspring,
  delays = delays,
  event_probs = event_prob_opts(
    asymptomatic = 0.5,
    presymptomatic_transmission = 0.15,
    symptomatic_traced = 0.5
  ),
  interventions = intervention_opts(quarantine = FALSE),
  sim = sim
)

set.seed(1)
outbreak_Q <- scenario_sim(
  n = nsim,
  initial_cases = initial_cases,
  offspring = offspring,
  delays = delays,
  event_probs = event_prob_opts(
    asymptomatic = 0.5,
    presymptomatic_transmission = 0.15,
    symptomatic_traced = 0.5
  ),
  interventions = intervention_opts(quarantine = TRUE),
  sim = sim
)

identical(outbreak_no_Q, outbreak_Q)
```

## Effect of quarantine in a COVID-like simulation

In the first example we will look at simulating a COVID-19 outbreak, following the parameterisation of @Hellewell2020.

```{r covid-params}
covid_offspring <- offspring_opts(
  community = \(n) rnbinom(n = n, mu = 2.5, size = 0.16),
  isolated = \(n) rnbinom(n = n, mu = 0, size = 1)
)
covid_delays <- delay_opts(
  incubation_period = \(n) rweibull(n = n, shape = 2.322737, scale = 6.492272),
  onset_to_isolation =\(n) rweibull(n = n , shape = 1.651524, scale = 2.305172)
)
covid_event_probs <- event_prob_opts(
  asymptomatic = 0.1, 
  presymptomatic_transmission = 0.15, 
  symptomatic_traced = 0.8
)

nsim <- 1000
```

```{r, run-covid-no-Q}
outbreak_no_Q <- scenario_sim(
  n = nsim, 
  initial_cases = initial_cases, 
  offspring = covid_offspring, 
  delays = covid_delays, 
  event_probs = covid_event_probs, 
  interventions = intervention_opts(quarantine = FALSE),
  sim = sim
)
```

From this outbreak we can look at the mean effective reproduction number ($\bar{R}_\text{eff}$) from each of the `r nsim` simulation replicates, as well as the number of generations (in terms of branching process generations) for each simulation replicate ($N_\mathrm{gen}$), and the final size of the outbreak ($Z$, at the maximum number of days (`r sim$cap_max_days`), or at the maximum number of cases (`r sim$cap_cases`)).

We define $\bar{R}_\text{eff}$ as the realised effective reproduction number per generation, averaged across all generations of a simulated outbreak. For each generation $g$ of an outbreak, the per-generation effective reproduction number is

$$
R_\text{eff}^{(g)} = \frac{c_g}{n_g},
$$

where $c_g$ is the number of secondary cases produced by infectors in generation $g$ and $n_g$ is the number of infectors in generation $g$. The mean effective reproduction number for the simulation is then

$$
\bar{R}_\text{eff} = \frac{1}{G} \sum_{g=1}^{G} R_\text{eff}^{(g)},
$$

where $G$ is the total number of generations of the outbreak.

```{r, Reff-no-Q}
# the R_eff is the same for all outbreak weeks so we get it from the last week
R_eff_no_Q <- outbreak_no_Q[week == max(week), effective_r0]
head(R_eff_no_Q)
```

```{r, n-gen-no-Q}
# the number of generations is the same for all outbreak weeks so we get it from the last week
n_gen_no_Q <- sapply(outbreak_no_Q[week == max(week), cases_per_gen], length)
head(n_gen_no_Q)
```

```{r, outbreak-size-no-Q}
# the cumulative number of cases in the outbreak from the last week
outbreak_size_no_Q <- outbreak_no_Q[week == max(week), cumulative]
head(outbreak_size_no_Q)
```

We can repeat the above simulation, but this time _activate quarantine_. 

```{r, run-covid-Q}
outbreak_Q <- scenario_sim(
  n = nsim, 
  initial_cases = initial_cases, 
  offspring = covid_offspring, 
  delays = covid_delays, 
  event_probs = covid_event_probs, 
  interventions = intervention_opts(quarantine = TRUE),
  sim = sim
)
```

We can again extract $\bar{R}_\text{eff}$, $N_\mathrm{gen}$ and the outbreak size.

```{r, Reff-Q}
# the R_eff is the same for all outbreak weeks so we get it from the last week
R_eff_Q <- outbreak_Q[week == max(week), effective_r0]
```

```{r n-gen-Q}
# the number of generations is the same for all outbreak weeks so we get it from the last week
n_gen_Q <- sapply(outbreak_Q[week == max(week), cases_per_gen], length)
```

```{r outbreak-size-Q}
# the cumulative number of cases in the outbreak from the last week
outbreak_size_Q <- outbreak_Q[week == max(week), cumulative]
```

And compare these by plotting them as either a density or histogram:

```{r compare-covid}
compare_Q <- data.table(
  R_eff = c(R_eff_no_Q, R_eff_Q),
  n_gen = c(n_gen_no_Q, n_gen_Q),
  outbreak_size = c(outbreak_size_no_Q, outbreak_size_Q),
  quarantine = c(rep(FALSE, length(R_eff_no_Q)), rep(TRUE, length(R_eff_Q)))
)
```

```{r plot-Reff-covid}
# kernel density bounded at zero so the curve does not extend below the support
dens_R_eff <- compare_Q[, {
  d <- density(R_eff, from = 0)
  .(x = d$x, y = d$y)
}, by = quarantine]

tinyplot(
  y ~ x | quarantine,
  data = dens_R_eff,
  type = "area",
  fill = "by",
  theme = "clean",
  xlab = expression(bar(R)[eff])
)
tinyplot_add(type = type_vline(v = 1), lty = 2)
```

```{r plot-n-gen-covid}
dens_n_gen <- compare_Q[, {
  d <- density(n_gen, from = 0)
  .(x = d$x, y = d$y)
}, by = quarantine]

tinyplot(
  y ~ x | quarantine,
  data = dens_n_gen,
  type = "area",
  fill = "by",
  theme = "clean",
  xlab = "Number of outbreak generations"
)
```

```{r plot-outbreak-size-covid}
# collapse all outbreaks sizes above cap 
compare_Q[, outbreak_size := pmin(outbreak_size, sim$cap_cases + 1L)]

tinyplot(
  ~ outbreak_size | quarantine,
  data = compare_Q,
  type = "hist",
  fill = "by",
  theme = "clean",
  alpha = 0.5,
  xlab = "Outbreak size (cumulative cases)"
)

```

```{r cap-rates-covid}
# percentage of simulation replicates that reached the case cap (cap_cases)
cap_rate_no_Q <- mean(outbreak_size_no_Q >= sim$cap_cases) * 100
cap_rate_Q    <- mean(outbreak_size_Q    >= sim$cap_cases) * 100
```

Across the three metrics, quarantine produces a small shift in the mean of $\bar{R}_\text{eff}$ but a substantial change in the _shape_ of all three distributions. In $\bar{R}_\text{eff}$, quarantine moves mass from a supercritical $\bar{R}_\text{eff}$ near 1.2 into an extinction mode near 0. The number of generations shows a similar redistribution: more outbreaks die out within fewer generations, and the right tail of long-running chains has less mass when quarantine is active. For outbreak size, under quarantine, more outbreaks are small ($<1,000$ total cases) and only `r sprintf("%.1f", cap_rate_Q)`% of replicates reach the case cap, whereas without quarantine `r sprintf("%.1f", cap_rate_no_Q)`% of replicates reach the maximum case cap.

::: {.alert .alert-info}

_Technical detail_

The mechanism behind this regime-switching is that quarantine lowers the `isolated_time` of symptomatic, traced infectees whose own onset would otherwise fall after their infector's isolation. This shifts the cut point in `sample_offspring()` such that fewer community offspring survive — a small per-infector reduction in realised reproduction number that compounds multiplicatively across generations and pushes a non-trivial fraction of chains across the stochastic extinction boundary. $\bar{R}_\text{eff}$ as a single number understates the effect because it averages across both regimes; the per-replicate distributions, and the downstream outbreak-size cap rate, reveal it.

:::

In the next sections we explore how this effect varies with the assumed onset-to-isolation delay distribution and with the proportions of asymptomatic cases and traced contacts.

## Varying onset-to-isolation delay distributions

When comparing the counterfactual scenarios of quarantine being inactive or active, the timing of symptom onset and the timing of isolation determine whether an infectee is removed from the community after a delay after their symptom onset time, or their infector's isolation time. Here we explore the impact the choice of onset-to-isolation delay has on the impact of quarantine on outbreak disease transmission.

We compare three onset-to-isolation distributional forms: a single value (i.e. Dirac Delta function), a uniform distribution, and a right-skewed distribution (Weibull distribution).

```{r, eval=FALSE, fig.show="hold", out.width="32%"}
dirac <- data.table(x = seq(0, 20, by = 0.05))
dirac[, y := ifelse(x == 10, 1, 0)]
tinyplot(
  y ~ x,
  data = dirac,
  type = "l", 
  lwd = 3,
  xlab = "Time (days)", 
  ylab = "Probability density",
  theme = "clean",
  main = "Dirac Delta Function"
)

unif <- data.table(x = seq(-5, 25, by = 0.05))
unif[, y := dunif(x, min = 0, max = 20)]
tinyplot(
  y ~ x,
  data = unif,
  type = "l",
  lwd = 3, 
  col = "darkgreen",
  xlab = "Time (days)",
  ylab = "Density",
  theme = "clean",
  main = "Uniform Distribution"
)

weibull <- data.table(x = seq(0, 20, by = 0.05))
weibull[, y := dweibull(x, shape = 1.7, scale = 5.6)]
tinyplot(
  y ~ x,
  data = weibull,
  type = "l",
  lwd = 3, 
  col = "brown",
  xlab = "Time (days)", 
  ylab = "Density",
  theme = "clean",
  main = "Weibull Distribution"
)
```

Now we compare the effect of quarantine for each of these onset-to-isolation parameterisations.

```{r, eval=FALSE}
delays <- list(
  dirac = delay_opts(
    incubation_period = \(n) rep(10, n),
    onset_to_isolation = \(n) rep(10, n)
  ),
  unif = delay_opts(
    incubation_period = \(n) runif(n = n, min = 0, max = 20), 
    onset_to_isolation = \(n) runif(n = n, min = 0, max = 20)
  ), 
  weibull = delay_opts(
    incubation_period = \(n) rweibull(n = n, shape = 1.7, scale = 5.6),
    onset_to_isolation = \(n) rweibull(n = n, shape = 1.7, scale = 5.6)
  )
)

params <- setDT(expand.grid(
  asymptomatic = c(0, 0.1),
  symptomatic_traced = c(1, 0.8),
  quarantine = c(FALSE, TRUE)
))

outbreak <- vector(mode = "list", length = length(delays) * nrow(params))
idx <- 1

for (i in seq_along(delays)) {
  for (j in seq_len(nrow(params))) {
    outbreak[[idx]] <- scenario_sim(
      n = nsim, 
      initial_cases = initial_cases, 
      offspring = offspring, 
      delays = delays[[i]],
      event_probs = event_prob_opts(
        asymptomatic = params[j, asymptomatic], 
        presymptomatic_transmission = 0.15, 
        symptomatic_traced = params[j, symptomatic_traced]
      ), 
      interventions = intervention_opts(quarantine = params[j, quarantine]),
      sim = sim
    )
    idx <- idx + 1
  }
}
```

We can again extract $\bar{R}_\text{eff}$ and $N_\mathrm{gen}$.

```{r, eval=FALSE}
# create parameter table for each delay type
pars <- rbindlist(list(params, params, params))
pars[, delay_type := c(rep("dirac", 8), rep("unif", 8), rep("weibull", 8))]

# the R_eff is the same for all outbreak weeks so we get it from the last week
R_eff <- lapply(outbreak, \(x) x[week == max(week), effective_r0])

# the number of generations is the same for all outbreak weeks so we get it from the last week
n_gen <- lapply(outbreak, \(x) sapply(x[week == max(week), cases_per_gen], length))

# the cumulative number of cases in the outbreak from the last week
outbreak_size <- lapply(outbreak, \(x) x[week == max(week), cumulative])

pars[, R_eff := list(R_eff)]
pars[, n_gen := list(n_gen)]
pars[, outbreak_size := list(outbreak_size)]

# unnest the list columns
cols_to_keep <- setdiff(names(pars), c("R_eff", "n_gen", "outbreak_size"))
compare_Q_dists <- pars[, .(R_eff = unlist(R_eff)), by = cols_to_keep]
compare_Q_dists[, n_gen := unlist(pars$n_gen)]
compare_Q_dists[, outbreak_size := unlist(pars$outbreak_size)]
```

We can plot just the scenarios that have complete contact tracing and no asymptomatic cases.

```{r, eval=FALSE}
dens_R_eff_full <- compare_Q_dists[
  asymptomatic == 0 & symptomatic_traced == 1,
  {
    d <- density(R_eff, from = 0)
    .(x = d$x, y = d$y)
  },
  by = .(quarantine, delay_type)
]

tinyplot(
  y ~ x | quarantine,
  data = dens_R_eff_full,
  type = "area",
  fill = "by",
  facet = ~delay_type,
  theme = "clean",
  xlab = expression(bar(R)[eff])
)
tinyplot_add(type = type_vline(v = 1), lty = 2)
```

```{r, eval=FALSE}
dens_n_gen_full <- compare_Q_dists[
  asymptomatic == 0 & symptomatic_traced == 1,
  {
    d <- density(n_gen, from = 0)
    .(x = d$x, y = d$y)
  },
  by = .(quarantine, delay_type)
]

tinyplot(
  y ~ x | quarantine,
  data = dens_n_gen_full,
  type = "area",
  fill = "by",
  facet = ~delay_type,
  xlab = "Number of outbreak generations",
  theme = "clean"
)
```

```{r, eval=FALSE}
tinyplot(
  ~ outbreak_size | quarantine, 
  data = compare_Q_dists[asymptomatic == 0 & symptomatic_traced == 1, ], 
  type = "hist", 
  fill = "by", 
  facet = ~delay_type,
  theme = "clean",
  alpha = 0.5
)
```

The $\bar{R}_\text{eff}$ plots show that quarantine has a small negative impact, but the more pronounced impact is on number of generations, which is reduced when quarantine is active.

---

At the start of this vignette we showed that when the probability of case being asymptomatic is one (i.e. no infections show symptoms) or the ascertainment of contacts by tracing is zero (i.e. no contacts are traced), both caused the effect of quarantine to go to zero. Next we repeat the above analysis, assuming three onset-to-isolation delay distributions, but with a non-zero value for the proportion of asymptomatic cases, and a non-one value for contacts ascertained. This will show how much of the quarantine effect shown in the two previous plots is diminished in these scenarios.

```{r, eval=FALSE}
dens_R_eff_partial <- compare_Q_dists[
  asymptomatic != 0 & symptomatic_traced != 1,
  {
    d <- density(R_eff, from = 0)
    .(x = d$x, y = d$y)
  },
  by = .(quarantine, delay_type)
]

tinyplot(
  y ~ x | quarantine,
  data = dens_R_eff_partial,
  type = "area",
  fill = "by",
  facet = ~delay_type,
  theme = "clean",
  xlab = expression(bar(R)[eff])
)
tinyplot_add(type = type_vline(v = 1), lty = 2)
```

```{r, eval=FALSE}
dens_n_gen_partial <- compare_Q_dists[
  asymptomatic != 0 & symptomatic_traced != 1,
  {
    d <- density(n_gen, from = 0)
    .(x = d$x, y = d$y)
  },
  by = .(quarantine, delay_type)
]

tinyplot(
  y ~ x | quarantine,
  data = dens_n_gen_partial,
  type = "area",
  fill = "by",
  facet = ~delay_type,
  xlab = "Number of outbreak generations",
  theme = "clean"
)
```

```{r, eval=FALSE}
tinyplot(
  ~ outbreak_size | quarantine, 
  data = compare_Q_dists[asymptomatic != 0 & symptomatic_traced != 1, ], 
  type = "hist", 
  fill = "by",
  facet = ~delay_type, 
  theme = "clean",
  alpha = 0.5
)
```

The plots show that the effect of quarantine on $\bar{R}_\text{eff}$ remains small, but the effect of quarantine on reducing the number of outbreak generations is diminished.

Investigating whether the change in $\bar{R}_\text{eff}$ and $N_\mathrm{gen}$ is due to the proportion of asymptomatics or contact tracing, or both, we change the model parameters across all possible combinations and show the results in the table below.

```{r, eval=FALSE}
compare_Q_summary <- compare_Q_dists[
  , .(
    mean_R_eff = mean(R_eff),
    median_R_eff = median(R_eff),
    mean_n_gen = mean(n_gen),
    median_n_gen = median(n_gen)
  ), 
  by = .(asymptomatic, symptomatic_traced, quarantine, delay_type)
]
```

```{r, eval=FALSE}
compare_Q_summary_w <- dcast(
  compare_Q_summary, 
  asymptomatic + symptomatic_traced + delay_type ~ quarantine, 
  value.var = c("mean_R_eff", "median_R_eff", "mean_n_gen", "median_n_gen")
)

compare_Q_summary_w[, `:=`(
  diff_mean_R_eff   = mean_R_eff_TRUE - mean_R_eff_FALSE,
  diff_median_R_eff = median_R_eff_TRUE - median_R_eff_FALSE,
  diff_mean_n_gen   = mean_n_gen_TRUE - mean_n_gen_FALSE,
  diff_median_n_gen = median_n_gen_TRUE - median_n_gen_FALSE
)]

res <- compare_Q_summary_w[, list(
  asymptomatic, 
  symptomatic_traced, 
  delay_type, 
  diff_mean_R_eff, 
  diff_median_R_eff, 
  diff_mean_n_gen,
  diff_median_n_gen
)]

# format column names
col_names <- c(
  "% Asym.", "% CT", "O2I Delay", "Mean $\\\\Delta \\\\bar{R}_\\text{eff}$",
  "Median $\\\\Delta \\\\bar{R}_\\text{eff}$", "Mean $\\\\Delta N_\\mathrm{gen}$",
  "Median $\\\\Delta N_\\mathrm{gen}$"
)

kable(res, col.names = col_names, row.names = FALSE, digits = 3)
```

The table shows the reduction in mean and median $\bar{R}_\text{eff}$ and $N_\mathrm{gen}$ caused by quarantine for each parameter set. If the diff is negative it means quarantine reduced the $\bar{R}_\text{eff}$ and the number of generations ($N_\mathrm{gen}$).

---

## Varying incubation period and onset-to-isolation delay distributions

Lastly we compare different values of the incubation period and onset-to-isolation to see how it affects the $\bar{R}_\text{eff}$ and $N_\mathrm{gen}$.


```{r, eval=FALSE}
# lseq <- function(from, to, length.out) {
#   exp(seq(log(from), log(to), length.out = length.out))
# }
# means <- lseq(from = 1, to = 50, length.out = 5)

means <- c(1, 5, 10)
params <- expand.grid(incub_mean = means, o2i_mean = means)

R_eff <- vector(mode = "list", length = nrow(params))
for (i in seq_len(nrow(params))) {
  for (j in c(FALSE, TRUE)) {
    cat("Running simulation ", i, "... Quarantine: ", j, "\n")
    incub <-  epiparameter::convert_summary_stats_to_params(
      "weibull", mean = params$incub_mean[i], sd = 2
    )
    o2i <-  epiparameter::convert_summary_stats_to_params(
      "weibull", mean = params$o2i_mean[i], sd = 2
    )
    out <- scenario_sim(
      n = nsim,
      initial_cases = 5,
      offspring = offspring_opts(
        community = \(n) rnbinom(n = n, mu = 2, size = 1),
        isolated = \(n) rnbinom(n = n, mu = 0, size = 1)
      ),
      delays = delay_opts(
        incubation_period = \(n) rweibull(n = n, shape = incub$shape, scale = incub$scale),
        onset_to_isolation = \(n) rweibull(n = n, shape = o2i$shape, scale = o2i$scale)
      ),
      event_probs = event_prob_opts(
        asymptomatic = 0.1,
        presymptomatic_transmission = 0.15,
        symptomatic_traced = 0.8
      ),
      interventions = intervention_opts(quarantine = j),
      sim = sim_opts(cap_max_days = 100, cap_cases = 5000)
    )
    if (isFALSE(j)) {
      # store non-quarantine R_eff
      R_eff[[i]] <- out[week == max(week), effective_r0]
    } else {
      # minus the quarantine R_eff
      R_eff[[i]] <- R_eff[[i]] - out[week == max(week), effective_r0]
    }
  }  
}

# head(R_eff)
R_eff_median <- sapply(R_eff, median)
head(R_eff_median)

params$R_eff <- R_eff_median
params

ggplot2::ggplot(data = params) +
  ggplot2::geom_tile(
    mapping = ggplot2::aes(x = o2i_mean, y = incub_mean, fill = R_eff)
  ) +
  ggplot2::scale_y_continuous(name = "Mean onset-to-isolation delay (days)") +
  ggplot2::scale_x_continuous(name = "Mean incubation period (days)") +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::labs(
    fill = expression(atop("Median reduction in ",  R[t] * "\n from quarantine"))
  )+
  ggplot2::theme_bw()
```

## References